Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics

### 2023-10-07 03_marlin_paper.qmd
R1 submitted

Title: Simulating Benefits, Costs, and Tradeoffs of Spatial Management in Marine
Social-Ecological Systems

@@

Designing effective spatial management strategies is challenging because marine
ecosystems are highly dynamic and opaque, and extractive entities such as
fishing fleets respond endogenously to ecosystem changes in ways that depend
upon ecological and policy context. We present a tool, `marlin`, that can be
used to efficiently simulate the bio-economic dynamics of marine systems in
support of both management and research. We demonstrate `marlin’s` capabilities
by focusing on two case studies on the conservation and food production impacts
of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna
fishery. In the coastal coral reef example, we explore how heterogeneity in
species distributions and fleet preferences can affect distributional outcomes
of MPAs. In the pelagic case study, we show how our model can be used to assess
the climate resilience of different MPA design strategies, as well as the
climate sensitivity of different fishing fleets. This paper demonstrates how
intermediate complexity simulation of coupled bio-economic dynamics can help
communities predict and potentially manage trade-offs between conservation,
fisheries yields, and distributional outcomes of management policies affected by
spatial bio-economic dynamics.

@@

1.  Introduction
2.  Methods
    1.  Model Summary
    2.  Fleet Dynamics
    3.  Population Model
    4.  Case Studies
3.  Results
    1.  Coral Reefs
    2.  Pelagic Systems
4.  Discussion
5.  Acknowledgements
6.  Data Availability Statement
7.  References

@@

Communities around the world are increasingly looking to spatial strategies for
managing marine ecosystems . For example, the "30x30" movement calls for
implementation of marine protected areas (MPAs), a form of spatial management,
across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and
references therein]. However , designing spatial management strategies to
achieve desired objectives -- which may include recovery and resilience of
overfished species, increased food production and economic well being, and
equitable distribution of benefits - is not straightforward. Different species
have different resilience to fishing, distributions in space and time, and value
to fishing fleets; fishing fleets themselves can have varying ranges of
incentives and reactions to policy structures. These dynamics pose challenges
even for single-species assessment and management strategies , which are only
amplified when we consider management policies specifically designed around
multiple species, fleets, and spatial features in the oceans [@field2006a].

@@

Here we present the core methods and functionality of `marlin`, and demonstrate
its use in two case-study applications; a coral reef system and a pelagic
system. Our results show how interacting ecological, economic, and design
attributes can affect the degree of benefits, costs, and trade offs between
conservation and food production outcomes of MPAs. More broadly, this paper both
demonstrates the functionality of our model and the critical need for
considering sufficiently realistic coupled ecological and economic dynamics in
policy evaluations.

@@

`marlin` follows a "models of intermediate complexity" philosophy
[@plagányi2014] targeted specifically towards capturing heterogeneous habitats,
movement dynamics, and the simultaneous impacts of fishing fleets across
multiple species (frequently called "technical interactions"). In its general
form, `marlin` simulates the behavior of populations of spatially explicit
biologically independent age-structured animal populations affected by one or
more fishing fleets in time steps and a spatial resolution specified by the
user.

@@

We define a few commonly used terms here. *Yields* refers to the volume of catch
from fishing activity. *Spawning stock biomass* refers to the total biomass of
reproductively mature fish in the water, a function of numbers, weight,
fecundity, and sexual maturity at age. We measure the size of the population by
the ratio of the *spawning stock biomass* (SSB) in a given time step relative to
*unfished spawning stock biomass* (SSB0), SSB/SSB0. So, a SSB/SSB0 value of 1
means that the population is unfished, a SSB/SSB0 value of 0 means the
population is extinct.

@@

1.  Trade-offs and distributional outcomes for food production and multi-species
    biomass in a coastal coral-reef fishery
2.  Implications of climate-driven range shifts for MPA design in a pelagic
    fishery

@@

These species can be caught by fishing fleets. A fishing fleet is defined in
`marlin` by a set of fishers that have the same fishing skill, prices, and
contact selectivity [@sampson2014] for individual species (each of which we
denote as a *metier*). For example, both a longline and purse seine fleet may
capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can
be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
Each fishing fleet distributes its fishing effort in space according to a
specified spatial allocation rule (see *Spatial Allocation of Effort* ), for
example in proportion to profit-per-unit-effort, conditional on management
policies in places such as quotas and/or the presence of any spatial
restrictions such as MPAs. It is important to note that contact selectivity
reflects the ability of the fishing method in question to capture fish of
different sizes that come into "contact" with the gear. The contact selectivity
of each of the specified *metiers* will then interact with the distribution of
fish sizes and fishing effort in space simulated by `marlin` to produce a net
"population" selectivity, which may differ from the individual contact
selectivities of each of the *metiers*.

@@

Users can use the model to explore general properties of spatially explicit
social-ecological systems, or can parameterize it to explore key uncertainties
of more tactical applications. It is very important note that the parameters of
this model cannot be fit to data directly within the `marlin` package; users
must set model parameters themselves based on externally available data and
their best judgement. On the biological side, at a minimum users must for each
species they wish to simulate supply the common or scientific name of the
species in question, a measure of the level of fishing intensity the species is
experiencing at the start of the simulation, and the diffusion rates for adult
and larvae. Given a common or scientific name, the model will then input default
life history parameters based on FishLife [internet connection required,
@thorson2020], though users are encouraged to check these values and supply
their own life history parameter values when possible. On the fleet side, users
must for each fleet they wish to simulate at a minimum specify the selectivity
curves for each species caught by the fleet, as well relative price per unit
weights greater than 0 for any target species. Species caught by but not
targeted by the fleet can be represented by prices of 0 (not targeted) or below
0 (actively avoided). We provide two case studies here designed to show users
how the `marlin` model can be used to address the kinds of questions
stakeholders might face if designing an MPA network for a coastal coral reef or
a pelagic system.

@@

Revenues for each fleet in a time step are a function of the total amount of
each species caught and its price ($\Pi$). The amount caught is a function of
the selectivity at age of each species for each fleet ($\alpha$), the fishing
efficiency of the fleet for that that species (also called "catchability", *q*),
the amount of fishing effort of the fleet in question in that patch in a given
time step (*E*), and the total instantaneous fishing mortality (*u*) at age
(*a*) (including all other fleets) for that particular species in that patch and
time step.

@@

Total catches or yield *Y* are calculated through the Baranov equation
[@baranov1918], which appropriately accounts for the total instantaneous
mortality (both fishing and natural. *z*) and divides the total amount of the
population biomass (*b*) killed between natural (*m*) and fishing (*u*) sources,
with the amount of biomass killed through fishing called "yield".

@@

Selectivity at age $\alpha$ is modeled through either a logistic form or a
dome-shaped form. The logistic form is based on the lengths *l* at which 50% of
individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which is
the difference between the length at 50% selectivity and the length at 95%
selectivity.

@@

Lastly *z* is total mortality, the sum of fishing mortality and natural
mortality (*m*).

@@

The degree of fishing pressure exerted by a given set of fishing fleets on each
species is a function of a range of parameters including the total amount of
effort (*E*), the selectivity ogives ($\alpha$,the curve specifying the
proportion of individuals at a given age or size that will be caught by a unit
of fishing mortality), fishing cost (*C*), the spatial distribution of the
species affected by the fleet, the relative prices across species ($\Pi$), and
the catchability coefficient of each fleet for each species ($q$). `marlin`
provides a tuning option to help users achieve desired biological outcomes from
their fleets.

@@

`marlin`'s movement dynamics are based on a continuous-time Markov chain (CTMC),
as described in @thorson2021a. Within this framework, the model allows for
movement to be broken down into three components of *advection* (drifting with
currents), *taxis* (active movement towards preferred habitat), and *diffusion*
(essentially remaining variation in movement not explained by advection or
taxis). For now, `marlin` focuses just on the diffusion and taxis components of
this model, assuming that advection is zero, though future extensions could
easily incorporate advection vectors from oceanographic models. In this way,
`marlin` allows users to run anything from a simple Gaussian dispersal kernel up
to a system governed by species that passively diffuse out from a core habitat
defined by a dynamic thermal range.

@@

The individual components (*M*) of the movement matrix ($\pmb{M}$) are filled
based on an adjacency matrix, which defines whether two patches are both
adjacent and water (as opposed to land or another physical barrier), a diffusion
rate $D$ defined in units of area of a patch per unit of time, and a habitat
preference function *H* in units of length of a side of a patch per unit time.
For example, if we are defining the time units as years and the distance units
as kilometers, for a tuna $D$ might be 1,000 $\frac{KM^2}{Year}$. We then use
parameters $\Delta_{t}$ and $\Delta_{d}$ parameters to translate the diffusion
rate $D$ to match the time step and patch size used in a simulation. For
example, if we were to run a model on a monthly timestep given time units of
years, then $\Delta_{t} = 1/12 years$. If one square patch in the simulation has
an area of 100km^2^, then $\Delta_d = 10KM$. This "scale free" parameterization
means that appropriate value of $D$ can be identified for a species and then
set, regardless of the time step or patch size used in the simulation model
itself. The taxis component of the movement process is a function of the
difference in habitat quality *H*. The habitat preference function itself can
take any form the user wishes. Exponentiating the difference in the habitat
preference function between patches turns the taxis matrix into a multiplier of
the diffusion rate *D*. As such, when creating habitat layers for simulation,
users can tune the scale of the habitat gradient function to result in realistic
multipliers of the diffusion rate. This parameterization ensure that the
off-diagonal elements of the movement matrix $\pmb{M}_{t,s,a}$ are all
non-negative, a requirement of the CTMC method.

@@

Spawning stock biomass *SSB* is calculated by converting age to mean length,
calculating weight at age, maturity at age, and then calculating spawning stock
biomass as the sum of spawning potential at age in a given time step, taking
into account the potential for hyperallometry in the manner of @marshall2021.
Age is converted to length through the von Bertalanffy growth equation given
parameters asymptotic length ($l_\infty$), growth (*k*) and theoretical age at
length zero ($a0$) assuming log-normally distributed variation *u* in the length
at age with CV $\sigma_s$.

@@

Fisheries management is often concerned with measuring stock status relative to
maximum sustainable yield (MSY) based reference points, though the exact level
of stock status relative to MSY reference points desired by management or
societal objectives may vary widely. MSY based reference points present a
problem for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of
fishery selectivity. Fishery selectivity in this model can vary by fleet, and
species can be distributed unevenly in both space and time. This means that the
net effective fishing selectivity on a species can vary depending on the
dynamics at a given moment, making the definition of an equilibrium concept such
as MSY challenging.

As such, we do not report MSY based reference points in the model by default,
with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which
tends to be conserved across selectivity patterns as it is primarily a function
of the mortality and flow of recruits relative to the maturity ogive. There are
many different strategies for estimating reference points in spatially explicit
systems [@kapur2021a]. We leave it to users to define and find relevant
reference point as required by their specific needs.

<!-- ### MPA Optimization -->

<!-- Our MPA optimization algorithm seems to create an MPA network that -->

<!-- maximizes an objective function, given by -->

<!-- $$ -->

<!-- obj_p = \Phi \times scale(\sum_f^F \pmb{\phi_{f,p}}) + (1 - \Phi)  scale(\pmb{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}}) -->

<!-- $$ {#eq-obj} -->

<!-- Where *scale* rescales the values to be between 0 and 1 such that the -->

<!-- weight parameter is comparable between vectors. The algorithm works by -->

<!-- first assigning every patch in the system to an MPA. For the first -->

<!-- round, the algorithm then calculates the equilibrium objective function -->

<!-- value resulting from sequentially removing each individual patch from -->

<!-- the MPA network. The cell whose removal from the MPA produces the -->

<!-- highest objective function value is then removed from the MPA network, -->

<!-- and the marginal value of the objective function each of the remaining -->

<!-- cells from the last steps are then stored. For the next steps, we -->

<!-- sampled 25% of the remaining MPA cells with probability proportional to -->

<!-- their marginal contribution to the objective function in the prior -->

<!-- iteration, selecting the cell whose removal produces the highest total -->

<!-- objective function as the cell to remove from the network, and updating -->

<!-- the marginal value of the sampled-but-not-selected cells. This in -->

<!-- essence is a sampling-importance-resampling algorithm. The algorithm -->

<!-- continues until no more cells are left in the MPA. This then provides -->

<!-- for a given objective weighting value $\Phi$ the near-optimal MPA -->

<!-- network at every MPA size from 0 to 100%. We say "near" as they is some -->

<!-- path dependency here, in that for example the network that maximizes the -->

<!-- objective function with 40% of the MPAs does not necessarily contain the -->

<!-- optimal network at 30%. However, the reversed nature of the algorithm, -->

<!-- starting at 100% MPA coverage and working better, produces a more stable -->

<!-- network with regards to this problem than starting at 0% and growing to -->

<!-- 100%. -->

@@

Each of the case studies contains too many parameters and options to be
succinctly presented in the text here, and we are concerned that attempting to
duplicate all of the parameters of the model within the text would simply
introduce errors and confusion. As such, and since these results are for
illustrative purposes only, readers should consult the accompanying code to view
the precise details of each simulation. Targeted tactical applications must
carefully consider and document all decisions made around model parameters.

@@

We make use of three potential "rule of thumb" MPA design strategies:

@@

3.  *Spawning Ground*: MPAs are centered on the grounds of a known spawning
    aggregation

@@

In our coastal coral reef example, we model the dynamics of four tropical
Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper
(*Lutjanus malabaricus*), a deep-water snapper (*Pristipomoides filamentosus*),
and a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers undergo
a mass migration to a spawning aggregation once per year, followed by the
sharks. Shallow-reef snappers stay in reefs closer to shore above a steep
drop-off year-round, while deep-water snappers stay in the deeper reefs past the
drop-off (@fig-coral).

These species are targeted by two different fleets. Fleet One primarily targets
the grouper and near-shore snapper populations, but will land any incidentally
captured sharks. Fleet One has a logistic selectivity pattern for all species,
as they retain any fish caught for consumption or sale. both grouper and
shallow-reef snapper due to demand for "plate sized" fish , and logistic
selectivity for sharks, as any shark caught is retained. Fleet One is totally
dependent on fishing for their livelihood, meaning the local community takes
advantage of every possible opportunity to fish, and as such we model it as a
"constant effort" fishery. Due to having less efficient boats, Fleet One has a
higher cost per distance coefficient than Fleet Two. Fleet One's home port is
located on the site of the grouper spawning grounds.

Fleet Two is a more commercial fleet that primarily targets the deep-water
snapper population. This fleet primarily sells their catch to local restaurants
and distributors where plate-sized fish are prized, and so for both snapper and
grouper Fleet Two has a dome-shaped selectivity pattern [@kindsvater2017]. While
plate-sized deep snapper are the primary target of Fleet Two, we model Fleet
Two's selectivity for deep snapper as logistic due to high levels of discard
mortality for deep-water snapper resulting from barotrauma. Fleet Two catches
groupers, though less than Fleet One, and receives no price for sharks due to
the requirements of a certification program through which they sell their
deep-water snapper. Accidental captures (bycatch) of sharks do occur, which
results in mortality. Fleet Two operates under open-access dynamics, as fishing
is not the only means of subsistence for this community; short-term effort
expands and contracts in response to profitability of the primarily
grouper-driven fishery. Fleet two's home port is located further north from
Fleet One.

We use `marlin` to simulate the outcomes for both food production and
conservation for each of the species and both of the fleets as a function of
both MPA size and MPA design strategy. For this exercise, MPAs are placed with
perfect information and have no design constraints for continuity (i.e. the
selected MPA design can be as patchwork as dictated by the data). We ran the
simulation for in quarter year time steps for 100 years ($\Delta_t = 1/4$) and
set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 144
patches for a total area of 2,000 KM^2^.

@@

#### Pelagic Ecosystems

@@

We model our pelagic use case loosely on the characteristics of the Western and
Central Pacific Ocean (WCPO) tuna fisheries. We simulate trajectories of
`r n_distinct(names(blue_fauna))` species commonly caught in the region,
including both the highly abundant skipjack tuna and the heavily depleted
oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use
publicly available data on catch-per-unit-effort of each of these species from
the WCPO as a very rough proxy for baseline habitat distributions, noting that
where possible, fishery-independent abundance indices would be preferable
(@fig-blue).

These pelagic species are caught by a longline fleet that primarily targets
large adult tunas such as bigeye and yellowfin for high-grade consumption, and a
purse seine fishery that primarily targets skipjack tunas for bulk canning.
Selectivities were modeled as logistic for the longline fleet, and dome-shaped
for the purse seine fleet, with selectivity values set to broadly match the
patterns reported in relevant [stock
assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
for each species. Both fleets operate under open-access dynamics with an effort
cap. The effort cap was set at the level of effort that resulted in the desired
levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue),
intended to simulate a scenario where managers step in to prevent further
expansion of fishing effort in a fully developed fishery. For
forward-simulation, open-access dynamics can result in effort decreasing in
response to profitability, but cannot result in effort beyond the effort cap set
for each fleet.

For this exercise, we focused on using `marlin` to assess resilience of the
selected Target Fishing and Rate "rule of thumb" MPA design strategies to a
climate-driven range shift. Specifically, we simulate an extreme example where
the centroid of each population shifts northward at a rate of 0.33 patches per
year, over a 20 year time horizon (@fig-blue-ssb). We designed MPA networks
given the conditions in the starting year, and then held that network constant
over the years of the experiment, running one simulation with and another
without the climate-drive range shift. We then compared the effects of this
range shift on food production and conservation outcomes from MPA networks
designed based on the pre-range shift world. We ran the pelagic simulation at a
quarterly level ($\Delta_t$ = 1/4 year), and set the area of each cell to be
roughly 97,000 KM^2^ across 144 patches each with a side length of roughly 311
KM, for a total area of 14,000,000 KM^2^, broadly commensurate with the area of
the WCPO.

@@

### Coral Reefs

@@

peak_coral_fleet_results |> filter(fleet == "fleet_two")

@@

### Pelagic Ecosystems

>>>>

Title: Simulating Benefits, Costs, and Trade-offs of Spatial Management in
Marine Social-Ecological Systems

@@

Designing effective spatial management strategies is challenging because marine
ecosystems are highly dynamic and opaque, and extractive entities such as
fishing fleets respond endogenously to ecosystem changes in ways that depend
upon ecological and policy context. We present a modeling framework, `marlin`,
that can be used to efficiently simulate the bio-economic dynamics of marine
systems in support of both management and research. We demonstrate `marlin’s`
capabilities by focusing on two case studies on the conservation and food
production impacts of marine protected areas (MPAs): a coastal coral reef and a
pelagic tuna fishery. In the coastal coral reef example, we explore how
heterogeneity in species distributions and fleet preferences can affect
distributional outcomes of MPAs. In the pelagic case study, we show how our
model can be used to assess the climate resilience of different MPA design
strategies, as well as the climate sensitivity of different fishing fleets. This
paper demonstrates how intermediate complexity simulation of coupled
bio-economic dynamics can help communities predict and potentially manage
trade-offs between conservation, fisheries yields, and distributional outcomes
of management policies affected by spatial bio-economic dynamics.

@@

1.  Introduction
2.  Methods
    1.  Model Summary
    2.  Fleet Dynamics
    3.  Population Model
    4.  Case Studies
3.  Results
    1.  Coral Reefs
    2.  Pelagic Systems
4.  Discussion
    1.  Insights from Case Studies
    2.  Putting `marlin` in Context
    3.  General Recommendations for Use
5.  Acknowledgements
6.  Data Availability Statement
7.  References

@@

Communities around the world are increasingly looking to spatial strategies for
managing marine ecosystems . For example, the "30x30" movement calls for
implementation of marine protected areas (MPAs), a form of spatial management,
across 30% of the world's oceans by the year 2030 [@grorud-colvert2021 and
references therein]. Recent agreements such as the United Nations (UN) agreement
on the conservation and sustainable use of marine biodiversity of areas beyond
national jurisdiction (BBNJ) calls for expanded area-based management in the
high seas. However, designing spatial management strategies to achieve desired
objectives -- which may include recovery and resilience of overfished species,
increased food production and economic well being, and equitable distribution of
benefits - is not straightforward. Different species have different resilience
to fishing, distributions in space and time, and value to fishing fleets;
fishing fleets themselves can have varying ranges of incentives and reactions to
policy structures. These dynamics pose challenges even for single-species
assessment and management strategies , which are only amplified when we consider
management policies specifically designed around multiple species, fleets, and
spatial features in the oceans [@field2006a].

@@

Here we present the core methods and functionality of `marlin`, and demonstrate
its use in two case-study applications; a coral reef system and a pelagic
system. Our results show how interacting ecological, economic, and design
attributes can affect the degree of benefits, costs, and trade-offs between
conservation and food production outcomes of MPAs. More broadly, this paper
demonstrates the functionality of our model and the critical need for
considering sufficiently realistic coupled ecological and economic dynamics in
policy evaluations.

@@

The model presented here focuses on representing the dynamics of heterogeneous
habitats and movement dynamics, along with the simultaneous impacts of fishing
fleets across multiple species (frequently called "technical interactions"). In
its general form, `marlin` simulates the behavior of populations of spatially
explicit biologically independent animal populations with age and subsequent
size structure affected by one or more fishing fleets in time steps and a
spatial resolution specified by the user.

@@

We define a few commonly used terms here. *Yields* refers to the volume of catch
from fishing activity. *Spawning stock biomass* refers to the total biomass of
reproductively mature fish in the water, a function of numbers, weight,
fecundity, and sexual maturity at age. We measure the size of the population by
the ratio of the *spawning stock biomass* (SSB) in a given time step relative to
*unfished spawning stock biomass* (SSB0), SSB/SSB0. An SSB/SSB0 value of 1 means
that the population is unfished, a SSB/SSB0 value of 0 means the population is
extinct.

@@

1.  Trade-offs and distributional outcomes for food production and multi-species
    biomass in a coastal coral reef fishery
2.  Implications of climate-driven range shifts for MPA design in a pelagic
    fishery

@@

These species can be caught by fishing fleets. A fishing fleet is defined in
`marlin` by a set of fishers that have the same fishing skill, prices, and
contact selectivity [@sampson2014] for individual species (each of which we
denote as a *metier*). For example, both a longline and purse seine fleet may
capture bigeye (*Thunnus obesus*) and skipjack tunas, but the longline fleet can
be made much more likely to catch larger bigeye than skipjack, and *vice versa*.
It is important to note that contact selectivity reflects the ability of the
fishing method in question to capture fish of different sizes that come into
"contact" with the gear. The contact selectivity of each of the specified
*metiers* will then interact with the distribution of fish sizes and fishing
effort in space simulated by `marlin` to produce a net "population" selectivity,
which may differ from the individual contact selectivities of each of the
*metiers* [@waterhouse2014; @sampson2014]. Each fishing fleet distributes its
fishing effort in space according to a specified spatial allocation rule (see
*Spatial Allocation of Effort* section), for example in proportion to
profit-per-unit-effort, conditional on management policies in places such as
quotas and/or the presence of any spatial restrictions such as MPAs.

@@

Users can use the model to explore general properties of spatially explicit
social-ecological systems, or can parameterize it to explore key uncertainties
of more tactical applications. It is very important note that the parameters of
this model cannot be fit to data directly within the `marlin` package; users
must set model parameters themselves based on externally available data and
their best judgement.

On the biological side, at a minimum users must for each species they wish to
simulate supply the common or scientific name of the species in question, a
measure of the level of fishing intensity the species is experiencing at the
start of the simulation, and the diffusion rates for adult and larvae. Given a
common or scientific name, the model will then input default life history
parameters based on FishLife [internet connection required, @thorson2020],
though users are encouraged to check these values and supply their own life
history parameter values when possible.

On the fleet side, users must for each fleet they wish to simulate at a minimum
specify the contact selectivity curves for each species caught by the fleet, as
well relative price per unit weights greater than 0 for any target species.
Species caught by but not targeted by the fleet can be represented by prices of
0 (not targeted) or below 0 (actively avoided).

@@

Revenues for each fleet in a time step are a function of the total amount of
each species caught and its price ($\Pi$). The amount caught is a function of
the contact selectivity at age of each species for each fleet ($\alpha$), the
fishing efficiency of the fleet for that that species (also called
"catchability", *q*), the amount of fishing effort of the fleet in question in
that patch in a given time step (*E*), and the total instantaneous fishing
mortality (*u*) at age (*a*) (including all other fleets) for that particular
species in that patch and time step.

@@

Total catches or yield *Y* are calculated through the Baranov equation
[@baranov1918], which accounts for the total instantaneous mortality (both
fishing and natural, *z*) and divides the total amount of the population biomass
(*b*) killed between natural (*m*) and fishing (*u*) sources, with the amount of
biomass killed through fishing called "yield".

@@

Contact selectivity at age $\alpha$ is modeled through either a logistic form or
a dome-shaped form. The logistic form is based on the lengths *l* at which 50%
of individuals are selected by the fishing gear, $l^{sel}$, and $\delta$ which
is the difference between the length at 50% selectivity and the length at 95%
selectivity.

@@

Lastly *z* is total mortality, the sum of fishing mortality (*u*) and natural
mortality (*m*).

@@

The degree of fishing pressure exerted by a given set of fishing fleets on each
species is a function of a range of parameters including the total amount of
effort (*E*), the contact selectivity ogives ($\alpha$,the curve specifying the
proportion of individuals at a given age or size that will be caught by a unit
of fishing mortality), fishing cost (*C*), the spatial distribution of the
species affected by the fleet, the relative prices across species ($\Pi$), and
the catchability coefficient of each fleet for each species ($q$). `marlin`
provides a tuning option to help users achieve desired biological outcomes from
their fleets.

@@

`marlin`'s movement dynamics are based on a continuous-time Markov chain (CTMC),
as described in @thorson2021a. Within this framework, the model allows for
movement to be broken down into three components of *advection* (drifting with
currents), *taxis* (active movement towards preferred habitat), and *diffusion*
(essentially remaining variation in movement not explained by advection or
taxis). For now, `marlin` focuses just on the diffusion and taxis components of
this model, assuming that advection is zero, though future extensions could
incorporate advection vectors from oceanographic models. In this way, `marlin`
allows users to run anything from a simple Gaussian dispersal kernel up to a
system governed by species that passively diffuse out from a core habitat
defined by a dynamic thermal range. The model currently focuses on diffusion and
taxis, which allows for general representation of animals following physical or
oceanographic features, but research on the incorporation and importance of
advective forces would be of value going forward.

@@

The individual components (*M*) of the movement matrix ($\pmb{M}$) are filled
based on an adjacency matrix, which defines whether two patches are both
adjacent and water (as opposed to land or another physical barrier), a diffusion
rate $D$ defined in units of area of a patch per unit of time, and a habitat
preference function *H* in units of length of a side of a patch per unit time.
For example, if we are defining the time units as years and the distance units
as kilometers, for a tuna $D$ might be 1,000 $\frac{KM^2}{Year}$. We then use
parameters $\Delta_{t}$ and $\Delta_{d}$ parameters to translate the diffusion
rate $D$ to match the time step and patch size used in a simulation. For
example, if we were to run a model on a monthly timestep given time units of
years, then $\Delta_{t} = 1/12 years$. If one square patch in the simulation has
an area of 100km^2^, then $\Delta_d = 10KM$. This "scale free" parameterization
means that appropriate value of $D$ can be identified for a species and then
set, regardless of the time step or patch size used in the simulation model
itself.

The taxis component of the movement process is a function of the difference in
habitat quality *H*. The habitat preference function itself can take any form
the user wishes. Exponentiating the difference in the habitat preference
function between patches turns the taxis matrix into a multiplier of the
diffusion rate *D*. As such, when creating habitat layers for simulation, users
can tune the scale of the habitat gradient function to result in realistic
multipliers of the diffusion rate. This parameterization ensures that the
off-diagonal elements of the movement matrix $\pmb{M}_{t,s,a}$ are all
non-negative, a requirement of the CTMC method.

@@

Spawning stock biomass *SSB* is calculated by converting age to mean length at
age, calculating weight at age, maturity at age, and then calculating spawning
stock biomass as the sum of spawning potential at age in a given time step,
taking into account the potential for hyperallometry in the manner of
@marshall2021. Age is converted to length through the von Bertalanffy growth
equation given parameters asymptotic length ($l_\infty$), growth (*k*) and
theoretical age at length zero ($a0$) assuming log-normally distributed
variation *u* in the length at age with CV $\sigma_s$.

@@

Fisheries management is often concerned with measuring stock status relative to
maximum sustainable yield (MSY) based reference points, though the exact level
of stock status relative to MSY reference points desired by management or
societal objectives may vary widely. MSY based reference points present a
problem for a multi-fleet and spatial-temporal model such as `marlin`. MSY and
the fishing mortality rate that would produce MSY, $F_{MSY}$, are a function of
fishery selectivity. Fishery selectivity in this model can vary by fleet, and
species can be distributed unevenly in both space and time. This means that the
net effective fishing selectivity on a species can vary depending on the
dynamics at a given moment, making the definition of an equilibrium concept such
as MSY challenging [@berger2017 and references therein].

As such, we do not report MSY based reference points in the model by default,
with the exception of the spawning stock biomass at MSY ($SSB_{MSY}$) which
tends to be better conserved across selectivity patterns as it is primarily a
function of the mortality and flow of recruits relative to the maturity ogive.
There are many different strategies for estimating reference points in spatially
explicit systems [@kapur2021a]. We leave it to users to define and find relevant
reference point as required by their specific needs.

@@

Each of the case studies contains too many parameters and options to be
succinctly presented in the text here, and we are concerned that attempting to
duplicate all of the parameters of the model within the text would simply
introduce errors and confusion. As such, and since these results are for
illustrative purposes only, readers should consult the accompanying code to view
the precise details of each simulation. Targeted applications must carefully
consider and document all decisions made around model parameters.

@@

We make use of three potential "rule of thumb" MPA design strategies in our case
studies

@@

3.  *Spawning Ground*: MPAs are centered on the grounds of a known spawning
    aggregation. This strategy is only used in the coral reef case study.

@@

In our coastal coral reef example, we model the dynamics of four tropical
Pacific species: a grouper (*Epinephelus fuscoguttatus*), a shallow-reef snapper
(*Lutjanus malabaricus*), a deep-reef snapper (*Pristipomoides filamentosus*),
and a reef shark (*Carcharhinus amblyrhynchos*). The simulated groupers undergo
a mass migration to a spawning aggregation once per year, followed by the
sharks. Shallow-reef snappers stay in reefs closer to shore above a steep
drop-off year-round, while deep-water snappers stay in the deeper reefs past the
drop-off (@fig-coral).

These species are targeted by two different fleets. Fleet One primarily targets
the grouper and near-shore snapper populations, but will land any incidentally
captured sharks. Fleet One has a logistic selectivity pattern for all species,
as they retain any fish caught for consumption or sale. Fleet One is totally
dependent on fishing for their livelihood, meaning the local community takes
advantage of every possible opportunity to fish, and as such we model it as a
"constant effort" fishery. Due to having less efficient boats, Fleet One has a
higher cost per distance coefficient than Fleet Two. Fleet One's home port is
located on near the site of the grouper spawning grounds.

Fleet Two is a more commercial fleet that primarily targets the snapper
populations. This fleet primarily sells their catch to local restaurants and
distributors where plate-sized fish are prized, and so for both snapper and
grouper Fleet Two has a dome-shaped selectivity pattern [@kindsvater2017]. While
plate-sized deep snapper are the primary target of Fleet Two, we model Fleet
Two's selectivity for deep snapper as logistic due to high levels of discard
mortality for deep-water snapper resulting from barotrauma. Fleet Two catches
groupers, though less than Fleet One, and receives no price for sharks due to
the requirements of a certification program through which they sell their
deep-water snapper. Accidental captures (bycatch) of sharks do occur, which
results in mortality. Fleet Two operates under open-access dynamics, as fishing
is not the only means of subsistence for this community; short-term effort
expands and contracts in response to profitability of the primarily
grouper-driven fishery. Fleet two's home port is located in the northwest corner
of the simulation space.

We used `marlin` to simulate the outcomes for both food production and
conservation for each of the species and both of the fleets as a function of
both MPA size and MPA design strategy. For this exercise, MPAs are placed with
perfect information and have no design constraints for continuity (i.e. the
selected MPA design can be as patchwork as dictated by the data). We ran the
simulation for in quarter year time steps for 20 years ($\Delta_t = 1/4$) and
set the area of each patch to be 5km^2^ ($\Delta_p = \sqrt{5}$), using 144
patches for a total area of 2,000 KM^2^.

@@

#### Pelagic Fishery

@@

We model our pelagic case study loosely on the characteristics of the Western
and Central Pacific Ocean (WCPO) tuna fisheries. We simulate trajectories of
`r n_distinct(names(blue_fauna))` species commonly caught in the region,
including both the highly abundant skipjack tuna and the heavily depleted
oceanic whitetip shark (*Carcharhinus longimanus*) (@fig-blue-ssb). We use
publicly available data on catch-per-unit-effort of each of these species from
the WCPO as a very rough proxy for baseline habitat distributions, noting that
where possible, fishery-independent abundance indices would be preferable
(@fig-blue).

These pelagic species are caught by a longline fleet that primarily targets
large adult tunas such as bigeye and yellowfin for high-grade consumption, and a
purse seine fishery that primarily targets skipjack tunas for bulk canning.
Contact selectivities were modeled as logistic for the longline fleet, and
dome-shaped for the purse seine fleet, with selectivity values set to broadly
match the patterns reported in relevant [stock
assessments](https://oceanfish.spc.int/en/ofpsection/sam/sam/210-stock-assessment-input-and-result-files-)
for each species. Both fleets operate under open-access dynamics with an effort
cap. The effort cap was set at the level of effort that resulted in the desired
levels of SSB/SSB0 for each species under open-access dynamics (@fig-blue),
intended to simulate a scenario where managers step in to prevent further
expansion of fishing effort in a fully developed fishery. For
forward-simulation, open-access dynamics can result in effort decreasing in
response to profitability, but cannot result in effort beyond the effort cap set
for each fleet.

For this exercise, we focused on using `marlin` to assess resilience of the
selected *Target Fishing* and *Rate* MPA design strategies to a climate-driven
range shift. Specifically, we simulate an extreme example where the centroid of
each population shifts northward at a rate of 0.2 patches per year over a 20
year time horizon (@fig-blue-ssb). We designed MPA networks given the conditions
in the starting year, and then held that network constant over the years of the
experiment, running one simulation with and another without the climate-drive
range shift. We then compared the effects of this range shift on food production
and conservation outcomes from MPA networks designed based on the pre-range
shift world. We ran the pelagic simulation at a quarterly level ($\Delta_t$ =
1/4 year), and set the area of each cell to be roughly 97,000 KM^2^ across 144
patches each with a side length of roughly 311 KM, for a total area of
14,000,000 KM^2^, broadly commensurate with the area of the WCPO.

@@

### Coastal Coral Reef Fishery

@@

@@

### Pelagic Fishery

Yields of the purse-seine fleet were more resilient to the range shift under the
*Rate* design strategy, while the longline fleet had the opposite result. Under
the *Rate* strategy, longline yields were relatively stable across a large range
of MPA sizes under the status quo conditions, but declined rapidly under the
range shift conditions. For the purse-seine fleet, yields were more stable under
the range shift conditions under the *Target Fishing* design scenario, but
declined quickly as a function of MPA size under the status quo conditions
(@fig-blue-results A). The *Target Fishing* strategy produced better yield
outcomes for both fishing fleets under the range shift conditions, but the
*Rate* strategy performed best under the status quo.
